---
title: "Coexpression analysis of yeast for gene prediction"
author: "Mark Ziemann"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    fig_width: 7
    fig_height: 7
theme: cosmo
---

## Intro


```{r,pkgs}

library("reshape2")
library(gplots)
library("corrr")

```

## Data download

```{r,dl1}

# accessions
if (!file.exists("scerevisiae_accessions.tsv.bz2")) {
  download.file("http://dee2.io/mx/scerevisiae_accessions.tsv.bz2",
    destfile="scerevisiae_accessions.tsv.bz2")
}

# qc
if (!file.exists("scerevisiae_qc.tsv.bz2")) {
  download.file("http://dee2.io/mx/scerevisiae_qc.tsv.bz2",
    destfile="scerevisiae_qc.tsv.bz2")
}

# gene level counts
if (!file.exists("scerevisiae_se.tsv.bz2")) {
  download.file("http://dee2.io/mx/scerevisiae_se.tsv.bz2",
    destfile="scerevisiae_se.tsv.bz2")
}


```

## GO

It looks like the server is not listening on the correct port.

```{r,go1}

# http://ftp.ebi.ac.uk/pub/databases/GO/goa/YEAST/README
# http://ftp.ebi.ac.uk/pub/databases/GO/goa/YEAST/

if (! file.exists("goa_yeast.gaf.gz") ) {
  download.file("http://ftp.ebi.ac.uk/pub/databases/GO/goa/YEAST/goa_yeast.gaf.gz",
    destfile="goa_yeast.gaf.gz")
}

if (! file.exists("goa_yeast.gpa.gz") ) {
  download.file("http://ftp.ebi.ac.uk/pub/databases/GO/goa/YEAST/goa_yeast.gpa.gz",
    destfile="goa_yeast.gpa.gz")
}

if (! file.exists("goa_yeast.gpi.gz") ) {
  download.file("http://ftp.ebi.ac.uk/pub/databases/GO/goa/YEAST/goa_yeast.gpi.gz",
    destfile="goa_yeast.gpi.gz")
}

if (! file.exists("sce_gene_info.tsv") ) {
  download.file("http://dee2.io/data/scerevisiae/sce_gene_info.tsv",
    destfile="sce_gene_info.tsv")
}

geneinfo <- read.table("sce_gene_info.tsv",header=TRUE)

gaf <- readLines("goa_yeast.gaf.gz")
gaf <- gaf[grep("!",gaf,invert=TRUE)]
gaf <- strsplit(gaf,"\t")
g <- sapply(gaf,"[[",3)
go <- sapply(gaf,"[[",5)
go <- cbind(g,go)
go <- unique(go)

# ontology must have 20 or more members
go <- go[which( go[,2] %in% names(which(table(go[,2])>=20)) ),]

```

## Filter

Here we filter for only runs that passed the QC.
It might be a good idea to remove lowly expressed genes as well.

```{r,filter1}

acc <- read.csv("scerevisiae_accessions.tsv.bz2",header=TRUE,sep="\t")
dim(acc)

qc <- read.csv("scerevisiae_qc.tsv.bz2",header=FALSE,sep="\t")
dim(qc)

se <- read.csv("scerevisiae_se.tsv.bz2",header=FALSE,sep="\t")
dim(se)

qcpass <- qc[which(qc$V3=="PASS"),1]
length(qcpass)

se <- se[which(se$V1 %in% qcpass),]
dim(se)

se <- as.matrix(acast(se, V1~V2, value.var="V3"))
dim(se)

```

## Aggregate SRA runs to experiments

```{r,agg1}

srx <- acc[,c(1,3)]
sesrx <- merge(srx,se,by.x="SRR_accession",by.y=0)
sesrx[,1] = NULL
sesrx <- aggregate(. ~ SRX_accession , sesrx , sum)
rownames(sesrx) <- sesrx[,1]
sesrx[,1] = NULL
se <- t(sesrx)
rm(sesrx,srx)

se[1:5,1:6]
dim(se)

```

## PCA analysis

No need to scale the data first.

```{r,pca1}

pc <- cmdscale(dist(t(se)),k=10)
barplot(colMeans(abs(pc)),names.arg=1:ncol(pc), main="screeplot on samples")
XMIN = min(pc[,1])*1.1
XMAX = max(pc[,1])*1.1
plot(pc[,c(1,2)] , xlab="PC1", ylab="PC2", type = "n" , xlim=c(XMIN,XMAX) , main="sample-wise PCA")
text(pc[,c(1,2)] , labels=colnames(se) )

ipc <- cmdscale(dist(se),k=10)
barplot(colMeans(abs(ipc)),names.arg=1:ncol(ipc), main="screeplot on genes")
XMIN = min(ipc[,1])*1.1
XMAX = max(ipc[,1])*1.1
plot(ipc[,c(1,2)] , xlab="PC1", ylab="PC2", type = "n" , xlim=c(XMIN,XMAX) , main="gene-wise PCA")
text(ipc[,c(1,2)] , labels=rownames(se) )

```

## Set average pseudosample

In this experiment, average the gene set expression values for the set members and the find
genes that correlate closely.
Hopefully the sample genes are identified and there might be some new ones too.

```{r,av1,fig.height=8,fig.width=8}

# make correlation matrix
se <- se[which(rowSums(se)!=0),]
nse <- se/colSums(se)*1000000
tnse_cor <- as.data.frame(correlate(t(nse)))
rownames(tnse_cor) <- tnse_cor[,1]
tnse_cor[,1] = NULL
hist(as.matrix(tnse_cor))
summary(as.vector(as.matrix(tnse_cor)))

```

```{r,av2}

# set up a list of GOs
gos <- unique(go[,2])

sets <- lapply(gos, function(x) { 
  myset <- go[which(go[,2] == x ),1]
  # convert from gene name to gene accession number
  myset <- geneinfo[which(geneinfo$GeneSymbol %in% myset ),1]
})

names(sets) <- gos

```

```{r,av3}

candidates <- lapply(sets[4], function(myset) {
  # extract data related to this set
  myset_nse <- nse[which(rownames(nse) %in% myset),]
  myset_tnse <- t(myset_nse)
  # calculate mean of the set
  #myset_mean <- rowMeans(myset_tnse)
  myset_mean <- apply(myset_tnse,1,median)
  myset_nsem <- rbind(myset_mean,myset_nse)
  myset_tnsem <- t(myset_nsem)
  myset_tnsem_cor <- as.data.frame(correlate(myset_tnsem))
  rownames(myset_tnsem_cor) <- myset_tnsem_cor[,1]
  myset_tnsem_cor[,1] = NULL
  #heatmap.2(as.matrix(myset_tnsem_cor),trace="none",scale="none")
  #hist(as.matrix(myset_tnsem_cor))
  #hist(as.matrix(myset_tnsem_cor[,1]))
  #par(mfrow=c(2,1))
  #hist(as.matrix(tnse_cor),xlim=c(-0.2,1))
  #tmp <- myset_tnsem_cor[,-1] 
  #tmp <- tmp[-1,] 
  #hist(as.matrix(tmp),xlim=c(-0.2,1))
  res <- apply(nse, 1, function(x) { cor(x,myset_mean) })
  # get the 3rd quartile
  setq3 <- summary(res[which(names(res) %in% myset)])[5]
  cand <- res[which(! names(res) %in% myset)]
  cand <- cand[which(cand > setq3)]
  return(cand)
})

names(candidates) <- gos

lapply(candidate, write, "candidates.txt", append=TRUE, ncolumns=10000)

```

## Set analysis

Calculate the PC distance between all genes and the ones in the set.

TODO:work out hypotenuse properly.

```{r,set1}

# start with a random one because GOA is down
plot(ipc[,c(1,2)] , xlab="PC1", ylab="PC2", pch=19 , xlim=c(XMIN,XMAX) , main="gene-wise PCA")
points( myset_ipc[,c(1,2)] , pch=19 , col="red"  )

myset_cor <- correlate(t(myset_ipc))
myset_cor[,1] = NULL
myset_cor <- as.matrix(myset_cor)
rownames(myset_cor) <- colnames(myset_cor)
hist(myset_cor)
summary(as.vector(myset_cor))

dat <- apply(ipc , 1 , function(y) {
  apply(myset_ipc , 1 , function(x) {
    sum(abs(y - x ))
  } )
} )

hist(apply(dat,2,median),xlim=c(0,1e6),breaks=1000)
summary(apply(dat,2,median))


```

## Saveimage

```{r,save}

save.image("coex.Rdata")

```

## Session info

```{r,sessioninfo}

sessionInfo()

```
